Linear and Differential Cryptanalysis of DES

### 2009-01-30 trunk/talk/draft1.tex
Spelling fixes and etc.
Starting conclusion

\begin{frame}
Let $P_L$ and $P_R$ be the left and right plaintext bits. Let $C_L$ and $C_R$ be the left and right ciphertext bits. Let $K_i$ be the round key for round $i$. \\
\vspace{5mm}
Using the S-Box 5 approximation and the Piling Up Principle we can combine the approximations to get
\begin{align*}
 P_L{25} & \oplus P_L{14} \oplus P_L{8} \oplus P_L{3} \oplus \\
         & C_L{25} \oplus C_L{14} \oplus C_L{8} \oplus C_L{3} \oplus \\
			& P_R{17} \oplus C_R{17} = K_1{26} \oplus K_3{26} \\
\end{align*}
The bias magnitude for this expression is for this expression is 
\[ (12/64)^2 + (1 - 12/64)^2 - 1/2 \approx 0.30 \]
This is the best expression for the 3-round DES.
\end{frame}

\begin{frame}
We perform the attack on the 4-round DES similiarly to how we performed the attack on 4-round SPN, by "undoing" the last round.\\
\vspace{5mm}
\pause
Note that in the last round, $C_L$ bits are not changed, only $C_R$ bits are affected. Note also that there is only one $C_R$ bit the 3-round equation. I
\vspace{5mm}
\pause
This bit, $C_R{17}$ is affected by the result of S-Box 1 in round 4. If we guess correctly for the 6 partial subkey bits $K_{4,1}$ to $K_{4,6}$, the 3-round equation will hold with high bias, otherwise it will likely hold with close to 0 bias.\\
\vspace{5mm}
\pause
Thus, this attack will give us $6$ of the $56$ key bits. What this means in terms of time saved over exhaustive search is this: If it takes a year to find a key through exhaustive search of $2^56$ bits, after this attack it will only take $5$ days to exhaustively search the remaining $2^50$ bits.
\end{frame}

@@

\subsection{Independence of S-Boxes}
\begin{frame}
Refresh math idea ... Independence assumption ...

>>>>

\begin{frame}
Let $P_L$ and $P_R$ be the left and right plaintext bits. Let $C_L$ and $C_R$ be the left and right ciphertext bits. Let $K_i$ be the round key for round $i$. \\
\vspace{5mm}
Using the S-Box 5 approximation and the Piling Up Principle we can combine the approximations to get
\begin{align*}
 P_L{25} & \oplus P_L{14} \oplus P_L{8} \oplus P_L{3} \oplus \\
         & C_L{25} \oplus C_L{14} \oplus C_L{8} \oplus C_L{3} \oplus \\
			& P_R{17} \oplus C_R{17} = K_1{26} \oplus K_3{26} \\
\end{align*}
The bias magnitude for this expression is
\[ (12/64)^2 + (1 - 12/64)^2 - 1/2 \approx 0.30 \]
This is the best expression for the 3-round DES.
\end{frame}

\begin{frame}
We perform the attack on the 4-round DES similiarly to how we performed the attack on 4-round SPN, by "undoing" the last round.\\
\vspace{5mm}
\pause
Note that in the last round, $C_L$ bits are not changed, only $C_R$ bits are affected. Note also that there is only one $C_R$ bit the 3-round equation.\\
\vspace{5mm}
\pause
This bit, $C_R{17}$ is affected by the result of S-Box 1 in round 4. If we guess correctly for the 6 partial subkey bits $K_{4,1}$ to $K_{4,6}$, the 3-round equation will hold with high bias, otherwise it will likely hold with close to 0 bias.\\
\vspace{5mm}
\pause
Thus, this attack will give us $6$ of the $56$ key bits. What this means in terms of time saved over exhaustive search is this: If it takes a year to find a key through exhaustive search of $2^{56}$ bits, after this attack it will only take $5$ days to exhaustively search the remaining $2^{50}$ bits.
\end{frame}

@@

\subsection{Independence of S-Boxes}
\begin{frame}
Note that throughout this discussion we have assumed that the S-Boxes were independant. This allowed us to use the Piling Up Principle to combine linear approximations of S-Boxes across rounds. \\
\vspace{3mm}
This assumption worked well for us in practice, but is not necessarily true. We could have proceeded differently without this assumption - [see John Manferdelli's Boolean Functions slides]
